---
title: "COVID-19 in Motion - World Data"
output:
  blogdown::html_page:
    dev: "svg"
author: "Houston Haynes"
date: '2018-11-08T14:59:43+05:30'
featuredpath: img
featured: COVID-19.png
ogfeatured: H3_og_wide_COVID-19.png
emblem: health.png
description: Looking at COVID data through time - first pass
draft: false
tags:
- COVID-19
categories: 
- Deep Dive
weight: 1
editor_options: 
  chunk_output_type: inline
---
```{r setup and file loads, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)


library(highcharter)
library(tidyverse)
library(lubridate)
library(widgetframe)
library(xts)
library(jsonlite)
library(fontawesome)

URL <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/owid-covid-data.csv"

coviddata <- read.csv(URL)

```

```{r Tweets from Faces of COVID, include=FALSE }

library(rtweet)
library(stringr)
library(dplyr)

appname <- "H3 Tech Blog"
consumer_key <- "Tb5d892Ozn7W3EP2uvm5RVQPH"
consumer_secret <- "i094J3ewdwxm1KKmY3n4hnJ5ygxwwrl2cjvcfYWCv4VjQwOHHh"
access_key <- "4434818952-MGJ5dPTsqmtm5sQHR3w2ebpDVt14wYsQeWBN9Bw"
access_secret <- "p5sI2H0Z8YeZ8Tiyvv8ivb6R6fYmPP7VgqoNkjEGgDBUm"

token <- create_token(
  appname, 
  consumer_key, 
  consumer_secret,
  access_key,
  access_secret)

COVIDtweets <- get_timeline("FacesOfCOVID", n = 1000, parse = TRUE, include_rts = FALSE, ) %>%
  filter(str_detect(text, 'died of COVID'))

COVIDtweets <- sample(COVIDtweets$status_id)


```

# Exercise in Scale

When I started working with the <a href="https://ourworldindata.org/" target="_blank">Our World In Data</a> COVID-19 data set I wanted to chart out a few things to get a *feel* for the data. I've done this work for quite a while and the first thing you often do is a form of *reality* check. In some cases I checked against other charts based on the same data and thought "that can't be right" only to plot it myself and see it directly. (looking at you, *Europe*) As we all have witnessed this year experts have been sounding the alarm since early on. Pandemics are like locomotives, they're difficult to stop once they gain momentum. The same goes for an effective response.

```{r New Cases by Continent, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE  }

library(zoo)
library(slider)

pop <- read.csv("../../data/population_by_country_2020.csv", header = TRUE)
data <- select(coviddata[with(coviddata, order(date, iso_code)),], 
               date, continent, location, new_cases)   %>%
  filter(location != "International", location != "World") %>%
  mutate(new_cases = ifelse(is.na(new_cases), 0, new_cases)) %>%
  mutate(date = ymd(date)) %>%
  left_join(pop, location = location) %>%
  select(date, continent, new_cases, population) %>%
  mutate(ncpm = ((new_cases / population) * 1000000)) %>%
  select (date, continent, new_cases, ncpm) %>%
  group_by(date, continent) %>%
  summarize(new_cases = sum(new_cases), ncpm = sum(ncpm)) %>%
  group_by(continent) %>%
  arrange(continent, date) %>%
  mutate(ncrm = slider::slide_dbl(new_cases, mean, .before = 3, .after = 3)) %>%
  mutate(ncpmrm = slider::slide_dbl(ncpm, mean, .before = 3, .after = 3)) %>%
  ungroup()

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

hchart(data, "areaspline", hcaes(x = date, y = new_cases, group = continent, labels = FALSE))%>%
  hc_plotOptions(series = list(stacking = 'normal')) %>%
  hc_title(text = "Global COVID Daily New Cases (raw)", align = "center") %>% 
  hc_add_theme(thm) %>%
  hc_subtitle(text = "Summed by continent :: Higher == BAD", align = "center")%>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE))

hchart(data, "areaspline", hcaes(x = date, y = ncrm, group = continent, labels = FALSE))%>%
  hc_plotOptions(series = list(stacking = 'normal')) %>%
  hc_title(text = "Global COVID Daily New Cases (smoothed)", align = "center") %>% 
  hc_add_theme(thm) %>%
  hc_subtitle(text = "Summed by continent :: Higher == BAD", align = "center")%>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE))

hchart(data, "areaspline", hcaes(x = date, y = ncpm, group = continent, labels = FALSE))%>%
  hc_plotOptions(series = list(stacking = 'normal')) %>%
  hc_title(text = "Global COVID Daily New Cases per Million (raw)", align = "center") %>% 
  hc_add_theme(thm) %>%
  hc_subtitle(text = "Summed by continent :: Higher == BAD", align = "center")%>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE))

hchart(data, "areaspline", hcaes(x = date, y = ncpmrm, group = continent, labels = FALSE))%>%
  hc_plotOptions(series = list(stacking = 'normal')) %>%
  hc_title(text = "Global COVID Daily New Cases per Million (smoothed)", align = "center") %>% 
  hc_add_theme(thm) %>%
  hc_subtitle(text = "Summed by continent :: Higher == BAD", align = "center")%>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE))

#frameWidget(widget,  width = "100%", height = "100%", options=frameOptions(allowfullscreen = T))


```

`r blogdown::shortcode("alert", "info", "Working with the Report:", "Use the legend to highlight, show or hide a series. Hovering over bands in the chart reveals detail for a given date.")`

So I took the data and plotted it on a map, to play on a day grain through time. The gradations are pretty subtle at the start, and that's because of the data getting *very* dark late in the range. 

```{r Global New Cases Per Million, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

data <- select(coviddata[with(coviddata, order(date, iso_code)),], date, iso_code, location, new_cases_smoothed_per_million)  %>%
  rename(value = new_cases_smoothed_per_million) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) 

ds <- data %>%
  group_by(iso_code) %>%
  do(item = list(
    iso_code = first(.$iso_code),
    sequence = .$value,
    value = first(.$value))) %>%
  .$item

#head(ds, 10)

vec <- sort(unique(pull(data, date)))

url <- "https://code.highcharts.com/mapdata/custom/world-robinson-highres.js"
tmpfile <- tempfile(fileext = ".json")
download.file(url, tmpfile)
geo <- readLines(tmpfile)
geo <- gsub(".* = ", "", geo)
map <- jsonlite::fromJSON(geo, simplifyVector = FALSE)

last_updated <- paste("Source: https://ourworldindata.org/  -  Report Last Updated:",  format(Sys.time(), "%a %b %d %Y %X"))

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Ubuntu")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

widget <- highchart(type = "map") %>% 
  hc_add_theme(thm) %>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_title(text = "Global COVID Daily New Cases per Million (smoothed)", align = "center") %>%
  hc_subtitle(text = "Press play button or use slider to change date", align = "center") %>%
  hc_add_series(data = ds,
                name = "New Cases per Million",
                mapData = map,
                joinBy = c("iso-a3", "iso_code"),
                borderWidth = 0.01) %>%
  hc_colorAxis (min = 0 ,
                max = 1000,
                stops = color_stops(colors = viridisLite::inferno(20, 
                                                                begin = 0.1, 
                                                                direction = -1))) %>%
  hc_motion (
    enabled = TRUE,
    startIndex = 23, 
    axisLabel = "date",
    labels = vec,
    series = 0,
    updateInterval = 10,
    playIcon = "fas fa-play",
    pauseIcon = "fas fa-pause",
    magnet = list(
      round = "floor",
      step = 0.1
      )
  ) %>%
  hc_legend(layout = "vertical", verticalAlign = "top",
            align = "right", valueDecimals = 0) %>% 
  hc_mapNavigation(
    enabled = TRUE,
    enableMouseWheelZoom = TRUE,
    enableDoubleClickZoom = TRUE
  ) %>%
  hc_credits(enabled = TRUE, 
             text = last_updated, 
             position = list(align = "left", x = 10, y = -5))


frameWidget(widget,  width = "100%", height = "40rem", options=frameOptions(allowfullscreen = T))

```

`r blogdown::shortcode("alert", "info", "Working with the Report:", "Aside from the play/pause and date slider, the map is *zoomable*. Hovering over a country will highlight and show the value for the current position on the timeline. Hovering while playing back will show updating changes to the values for that country.")`

`r blogdown::shortcode("alert", "warning", "Bug Report:", "The play/pause button for the map charts *works* but the icons are currently not visible. A fix is being researched now. Thanks for understanding.")`

# Daily COVID Mortality

One of the things I expected was to see some "chop" in early data as reporting became more consistent. Here I was surprised to see irregular bumps in the "smoothed" data. Once South America is removed from the set then things look much less strange. Such as it is, the visuals tell a convincing story of where things are going as of November, and it's *not* good.
```{r New Deaths by Continent, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE  }

library(dplyr)

data <- select(coviddata[with(coviddata, order(date, iso_code)),], date, continent, new_deaths_smoothed_per_million)   %>%
  rename(value = new_deaths_smoothed_per_million) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) %>%
  group_by(date, continent) %>%
  summarise_each(funs(mean))

data <- subset(data, continent != "") 

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu',color ='black'))))

widget <- hchart(data, "area", hcaes(x = date, y = value, group = continent))%>%
  hc_plotOptions(series = list(stacking = 'normal')) %>%
  hc_title(text = "Global COVID Daily New Deaths per Million (smoothed)", align = "center") %>% 
  hc_add_theme(thm) %>%
  hc_subtitle(text = "Summed by continent :: Higher == BAD", align = "center")%>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60')  %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE))

frameWidget(widget,  width = "100%", height = "100%", options=frameOptions(allowfullscreen = T))


```

Again here - like the area chart - the direction of what's happening in Europe is unmistakable. In fact it bothers me to the point that I'll probably end up sourcing some other data just so I can sleep tonight. But one thing we can see and possibly discard is the sudden jump in values in South America. Per our area chart above which showed that as the offending portion of the data set that was skewing the result.
 
```{r Global New Deaths Per Million, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

data <- select(coviddata[with(coviddata, order(date, iso_code)),], date, iso_code, location, new_deaths_smoothed_per_million)  %>%
  rename(value = new_deaths_smoothed_per_million) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) 

ds <- data %>%
  group_by(iso_code) %>%
  do(item = list(
    iso_code = first(.$iso_code),
    sequence = .$value,
    value = first(.$value))) %>%
  .$item

#head(ds, 10)

vec <- sort(unique(pull(data, date)))

url <- "https://code.highcharts.com/mapdata/custom/world-robinson-highres.js"
tmpfile <- tempfile(fileext = ".json")
download.file(url, tmpfile)
geo <- readLines(tmpfile)
geo <- gsub(".* = ", "", geo)
map <- jsonlite::fromJSON(geo, simplifyVector = FALSE)

last_updated <- paste("Source: https://ourworldindata.org/  -  Report Last Updated:",  format(Sys.time(), "%a %b %d %Y %X"))

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Ubuntu")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu',color ='black'))))

widget <- highchart(type = "map") %>% 
  hc_add_theme(thm) %>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_title(text = "Global COVID Daily New Deaths per Million (smoothed)", align = "center") %>%
  hc_subtitle(text = "Press play button or use slider to change date", align = "center") %>%
  hc_add_series(data = ds,
                name = "New Cases per Million",
                mapData = map,
                joinBy = c("iso-a3", "iso_code"),
                borderWidth = 0.01) %>%
  hc_colorAxis (min = 0 ,
                max = 20,
                stops = color_stops(colors = viridisLite::inferno(20, 
                                                                begin = 0.1, 
                                                                direction = -1))) %>%
  hc_motion (
    enabled = TRUE,
    startIndex = 23, 
    axisLabel = "date",
    labels = vec,
    series = 0,
    updateInterval = 10,
    magnet = list(
      round = "floor",
      step = 0.1
      )
  ) %>%
  hc_legend(layout = "vertical", verticalAlign = "top",
            align = "right", valueDecimals = 0) %>% 
  hc_mapNavigation(
    enabled = TRUE,
    enableMouseWheelZoom = TRUE,
    enableDoubleClickZoom = TRUE
  ) %>%
  hc_credits(enabled = TRUE, 
             text = last_updated, 
             position = list(align = "left", x = 10, y = -5))


frameWidget(widget,  width = "100%", height = "40rem", options=frameOptions(allowfullscreen = T))

```

<div class="separator">Faces of COVID-19</div>
<figure>
<div class="row">
  <div class="col-md-4">`r blogdown::shortcode("tweet", COVIDtweets[1])`</div>
  <div class="col-md-4">`r blogdown::shortcode("tweet", COVIDtweets[2])`</div>
  <div class="col-md-4">`r blogdown::shortcode("tweet", COVIDtweets[3])`</div>
</div>
<figcaption class="caption">The pandemic-related posts on this site are about more than data. <strong>Behind every number is a person, a family and a community.</strong> As reports are refreshed, new selections will also be chosen at random. To see how this is done, <a href="/post/faces-of-covid">see this sidebar</a>.</figcaption>
<hr>
</figure>

# Testing, Testing, 1, 1, 3...

While spelunking through the data I thought I'd take a look at testing. The good news, the numbers are increasing. 

```{r New Tests by Continent, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE  }

library(dplyr)

data <- select(coviddata[with(coviddata, order(date, iso_code)),], date, continent, new_tests_smoothed_per_thousand)   %>%
  rename(value = new_tests_smoothed_per_thousand) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) %>%
  group_by(date, continent) %>%
  summarise_each(funs(mean))

data <- subset(data, continent != "") 

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu',color ='black'))))

widget <- hchart(data, "area", hcaes(x = date, y = value, group = continent))%>%
  hc_plotOptions(series = list(stacking = 'normal')) %>%
  hc_title(text = "Global COVID Daily Tests per Thousand (smoothed)", align = "center") %>% 
  hc_add_theme(thm) %>%
  hc_subtitle(text = "Summed by continent :: Higher == GOOD", align = "center")%>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60')  %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE))

frameWidget(widget,  width = "100%", height = "100%", options=frameOptions(allowfullscreen = T))


```

Here's the flip-side of the previous two reports - where the numbers go *green* as they go higher. But again, it runs the risk of painting an inaccurate picture to show  less than 5 tests per 1,000 people as a positive indication. That value should be much, much higher - but in this case relative values matter. 

```{r Global New Test Per Thousand, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

data <- select(coviddata[with(coviddata, order(date, iso_code)),], 
               date, iso_code, location, new_tests_smoothed_per_thousand)  %>%
  rename(value = new_tests_smoothed_per_thousand) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) 

ds <- data %>%
  group_by(iso_code) %>%
  do(item = list(
    iso_code = first(.$iso_code),
    sequence = .$value,
    value = first(.$value))) %>%
  .$item

#head(ds, 10)

vec <- sort(unique(pull(data, date)))

url <- "https://code.highcharts.com/mapdata/custom/world-robinson-highres.js"
tmpfile <- tempfile(fileext = ".json")
download.file(url, tmpfile)
geo <- readLines(tmpfile)
geo <- gsub(".* = ", "", geo)
map <- jsonlite::fromJSON(geo, simplifyVector = FALSE)

last_updated <- paste("Source: https://ourworldindata.org/  -  Report Last Updated:", 
                      format(Sys.time(), "%a %b %d %Y %X"))

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Ubuntu")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu',color ='black'))))

widget <- highchart(type = "map") %>% 
  hc_add_theme(thm) %>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_title(text = "Global COVID Daily New Tests per Thousand (smoothed)", align = "center") %>%
  hc_subtitle(text = "Press play button or use slider to change date", align = "center") %>%
  hc_add_series(data = ds,
                name = "New Cases per Million",
                mapData = map,
                joinBy = c("iso-a3", "iso_code"),
                borderWidth = 0.01) %>%
  hc_colorAxis (min = 0 ,
                max = 15,
                stops = color_stops(colors = viridisLite::viridis(20, 
                                                                begin = 0.1, 
                                                                direction = -1))) %>%
  hc_motion (
    enabled = TRUE,
    startIndex = 23, 
    axisLabel = "date",
    labels = vec,
    series = 0,
    updateInterval = 10,
    playIcon = "fas fa-play",
    pauseIcon = "fas fa-pause",
    magnet = list(
      round = "floor",
      step = 0.1
      )
  ) %>%
  hc_legend(layout = "vertical", verticalAlign = "top",
            align = "right", valueDecimals = 0) %>% 
  hc_mapNavigation(
    enabled = TRUE,
    enableMouseWheelZoom = TRUE,
    enableDoubleClickZoom = TRUE
  ) %>%
  hc_credits(enabled = TRUE, 
             text = last_updated, 
             position = list(align = "left", x = 10, y = -5))


frameWidget(widget,  width = "100%", height = "40rem", options=frameOptions(allowfullscreen = T))

```

## How the resports were built

The area charts are relatively straightforward. The only data munging task was summing the values by continent and date. After that it was mainly about formatting and some tweaks to keep the display clean. 

<button class="btn btn-outline-dark btn-sm mt-2 mb-2" data-toggle="collapse" data-target="#code1">View the Code behind the First Area Chart</button>

<pre id="code1" class="collapse"><code class="r"># begin setup code chunk
library(highcharter)
library(widgetframe)
library(tidyverse)
library(lubridate)
library(xts)

URL <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/owid-covid-data.csv"

coviddata <- read.csv(URL)

# end setup code chunk

data <- select(coviddata[with(coviddata, order(date, iso_code)),], 
                date, continent, new_cases_smoothed_per_million)   %>%
  rename(value = new_cases_smoothed_per_million) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) %>%
  group_by(date, continent) %>%
  summarise_each(funs(mean))

data <- subset(data, continent != "") 

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

widget <- hchart(data, "areaspline", hcaes(x = date, y = value, group = continent, labels = FALSE))%>%
  hc_plotOptions(series = list(stacking = 'normal')) %>%
  hc_title(text = "Global COVID Daily New Cases per Million (smoothed)", align = "center") %>% 
  hc_add_theme(thm) %>%
  hc_subtitle(text = "Summed by continent :: Higher == BAD", align = "center")%>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE))

frameWidget(widget,  width = "100%", height = "100%", options=frameOptions(allowfullscreen = T))</code></pre>

The map charts were more complex. Aside from tying together the geojson map to the data (by joining iso-a3 in the map data to the iso_code value from the source data) there was building the sequence of data itself. This wasn't immediately apparent but <a href="https://jkunst.com/blog/posts/2016-04-12-adding-motion-to-choropleths/" target="_blank">this tutorial</a> was a great help.

<button class="btn btn-outline-dark btn-sm mt-2 mb-2" data-toggle="collapse" data-target="#code2">View the Code behind the First Map Chart</button>

<pre id="code2" class="collapse"><code class="r"># begin setup code chunk
library(highcharter)
library(widgetframe)
library(tidyverse)
library(lubridate)
library(xts)

URL <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/owid-covid-data.csv"

coviddata <- read.csv(URL)

# end setup code chunk

data <- select(coviddata[with(coviddata, order(date, iso_code)),], 
                date, iso_code, location, new_cases_smoothed_per_million)  %>%
  rename(value = new_cases_smoothed_per_million) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) 

ds <- data %>%
  group_by(iso_code) %>%
  do(item = list(
    iso_code = first(.$iso_code),
    sequence = .$value,
    value = first(.$value))) %>%
  .$item

vec <- sort(unique(pull(data, date)))

url <- "https://code.highcharts.com/mapdata/custom/world-robinson-highres.js"
tmpfile <- tempfile(fileext = ".json")
download.file(url, tmpfile)
geo <- readLines(tmpfile)
geo <- gsub(".* = ", "", geo)
map <- jsonlite::fromJSON(geo, simplifyVector = FALSE)

last_updated <- paste("Source: https://ourworldindata.org/  -  Report Last Updated:",
                        format(Sys.time(), "%a %b %d %Y %X"))

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Ubuntu")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

widget <- highchart(type = "map") %>% 
  hc_add_theme(thm) %>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_title(text = "Global COVID Daily New Cases per Million (smoothed)", align = "center") %>%
  hc_subtitle(text = "Press play button or use slider to change date", align = "center") %>%
  hc_add_series(data = ds,
                name = "New Cases per Million",
                mapData = map,
                joinBy = c("iso-a3", "iso_code"),
                borderWidth = 0.01) %>%
  hc_colorAxis (min = 0 ,
                max = 1000,
                stops = color_stops(colors = viridisLite::inferno(20, 
                                                                begin = 0.1, 
                                                                direction = -1))) %>%
  hc_motion (
    enabled = TRUE,
    startIndex = 23, 
    axisLabel = "date",
    labels = vec,
    series = 0,
    updateInterval = 10,
    playIcon = "fas fa-play",
    pauseIcon = "fas fa-pause",
    magnet = list(
      round = "floor",
      step = 0.1
      )
  ) %>%
  hc_legend(layout = "vertical", verticalAlign = "top",
            align = "right", valueDecimals = 0) %>% 
  hc_mapNavigation(
    enabled = TRUE,
    enableMouseWheelZoom = TRUE,
    enableDoubleClickZoom = TRUE
  ) %>%
  hc_credits(enabled = TRUE, 
             text = last_updated, 
             position = list(align = "left", x = 10, y = -5))


frameWidget(widget,  width = "100%", height = "40rem", options=frameOptions(allowfullscreen = T))</code></pre>

# Further analysis

Aside from drilling into other categories of data here and checking against other data sets, I'll also be looking at various forecast models and lag indicators. There will also be *drill down* into national and regional level data. In any case, as the story continues to unfold we should all brace for a very *dark* winter. 
