---
title: "COVID in India - A Perfect Storm?"
output:
  blogdown::html_page:
    dev: "svg"
author: "Houston Haynes"
date: '2021-05-23T14:59:43+05:30'
featuredpath: img
featured: COVID-19.png
ogfeatured: h3_og_wide_covid19.png
emblem: health.png
description: Confounding Congregate Events and Strains of the Virus Straining Systems to the Breaking Point
draft: false
weight: 2
tags:
- COVID-19
- R
- highcharts
categories: 
- Deep Dive
editor_options: 
  chunk_output_type: inline
---
```{r setup and file loads, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(highcharter)
library(tidyverse)
library(lubridate)
library(widgetframe)
library(xts)
library(jsonlite)

URL <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/owid-covid-data.csv"

coviddata <- read.csv(URL)

```

```{r Tweets_from_Faces_of_COVID, include=FALSE }

library(rtweet)
library(stringr)
library(dplyr)
library(DT)

appname <- "H3 Tech Blog"
consumer_key <- "Tb5d892Ozn7W3EP2uvm5RVQPH"
consumer_secret <- "i094J3ewdwxm1KKmY3n4hnJ5ygxwwrl2cjvcfYWCv4VjQwOHHh"
access_key <- "4434818952-MGJ5dPTsqmtm5sQHR3w2ebpDVt14wYsQeWBN9Bw"
access_secret <- "p5sI2H0Z8YeZ8Tiyvv8ivb6R6fYmPP7VgqoNkjEGgDBUm"

token <- create_token(
  appname, 
  consumer_key, 
  consumer_secret,
  access_key,
  access_secret)

COVIDtweets <- get_timeline("CTZebra", n = 1000, parse = TRUE, include_rts = FALSE, ) %>%
  filter(str_detect(text, 'died of #covid19'))

COVIDtweets <- sample(COVIDtweets$status_id)

# glimpse(COVIDtweets)

# IndiaCOVIDtweets <- search_tweets("india, doctor OR nurse OR 'healthcare worker', 'died of covid'", n = 1000, parse = TRUE, include_rts = FALSE, ) 

# IndiaCOVIDtweets <- sample(IndiaCOVIDtweets$status_id)

# glimpse(IndiaCOVIDtweets)

```

# My Cassandra Syndrome

From the early days of lock down last year I was just as worried for *a possible* COVID-20 and COVID-21 as *the current* COVID-19. And when reconnecting with friends and former co-workers, many of whom work in public health and study of infectious diseases, the collective sentiment was largely similar. A close friend and I puzzled over India, as he's from there and was *convinced* that it would see an outsized infection rate due to its population density. But we were still puzzled why India's big wave emerged so recently and so asymmetrically. Part of it is our shared suspicion that there was under-reporting of cases, but we both thought there was more to the story.

## Congregate Events

Taking the queue from my entries about contagion spreading due to end-of-year congregate events in `r blogdown::shortcode("url-link", "Canada", "/post/COVID-19-Canadian-Case-Study", "_self")` and `r blogdown::shortcode("url-link", "Israel", "/post/COVID-19-Israel-Case-Study", "_self")`, my friend supposed that the same would happen in India with the advent of the national `r blogdown::shortcode("url-link", "Kumbh Festival", "https://www.bbc.com/news/world-asia-india-56770460")`. Even though many of those public events are held outdoors, my friend explained some of the cultural norms had a great deal of close contact over extended periods. So I decided to dive back into the "Our World In Data" repository and line up India's infection data with the dates of the festival.

`r blogdown::shortcode("fancybox-iframe", "/post/COVID-19-India-a-perfect-storm_files/figure-html/widgets/widget_India_COVID_Cases_per_Million.html")`

```{r India_COVID_Cases_per_Million, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

data <- select(coviddata, date, location, new_cases_smoothed_per_million) %>%
  filter(date >= "2021-01-01") %>%
  rename(value = new_cases_smoothed_per_million) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) 

data.wide <-  pivot_wider(data, names_from = location, values_from = value)

data_xts <- xts(data.wide[,-1], order.by = data.wide$date)

last_updated <- paste("Source: https://ourworldindata.org/  -  Report Last Updated:",  format(Sys.time(), "%a %b %d %Y %X"))

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

widget <- highchart(type = "chart") %>%
  hc_add_theme(thm) %>%
  hc_title(text = "India's Kumbh Festival Dates Compared To COVID Infection Rates (per Million)", align = "center") %>%
    hc_chart(
      borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
  hc_plotOptions(series = list(marker = list(enabled = FALSE))) %>%
  hc_credits(enabled = TRUE, 
             text = last_updated, 
             position = list(align = "left", x = 10, y = -5)) %>%
  hc_add_series(data_xts$India, type = "area", name = "India", color = "#FF9933", showInLegend = FALSE) %>%
  hc_tooltip(valueDecimals = 2) %>%
  hc_xAxis(type = "datetime",
    plotLines = list(list(
      label = list(text = "First Bath: Maka Sankranti"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-01-14', tz = 'UTC'))
    ),
    list(
      label = list(text = "Second Bath: Mauni Amavasya"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-02-11', tz = 'UTC'))
    ),
    list(
      label = list(text = "Third Bath: Vasant Panchami"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-02-16', tz = 'UTC'))
    ),
    list(
      label = list(text = "Fourth Bath: Ram Navami"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-04-21', tz = 'UTC'))
    ),
    list(
      label = list(text = "First Shahi Bath: Maha Shivratri"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-03-11', tz = 'UTC'))
    ),
    list(
      label = list(text = "Second Shahi Bath: Somvati Amamvasya"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-04-12', tz = 'UTC'))
    )
    )) 

frameWidget(widget, width="100%", height="26rem")

```

I was fully expecting for each date to be followed by a "bump" in case rates similar to what occurred in Canada and Israel. And while it *can* be influenced by under-reporting it's hard to see *that* many events remaining flat. So what could be the reason for such a huge wave of cases? It's entirely possible that there's no correlation whatsoever, or that there's something tangential about the later events that were minor contributors. There are too many *unknown unknowns* for this to be an __aha!__ moment. But with news of more virulent strains migrating to and gaining a footing in India I thought I'd take a look at genomics records to see if there's anything *else* that emerges. So I dug into `r blogdown::shortcode("url-link", "Nextstrain's data", "https://nextstrain.org/ncov/global?c=pango_lineage")`. Even if there weren't any answers I was hoping it would lead to better questions.

`r blogdown::shortcode("fancybox-iframe", "/post/COVID-19-India-a-perfect-storm_files/figure-html/widgets/widget_India_COVID_Variants_By_Month.html")`

```{r India_COVID_Variants_By_Month, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

asia <- "/repo/next-strain-ncov/data/asia.csv"

asiadata <- read.csv(asia)

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

data <- select(asiadata, Collection.Data, Country, PANGO.Lineage, Clade) %>%
  filter(Country == "India" 
#         | Country == "Malaysia" 
#         | Country == "Bangladesh" 
#         | Country == "Nepal"
#         | Country == "Sri Lanka"
#         | Country == "Afghanistan"
#         | Country == "Pakistan"
#         | Country == "Maldives"
         ) %>%
  add_column(value = 1) %>%
  rename(Date = Collection.Data) %>%
  rename(Variant = PANGO.Lineage) %>%
  mutate(Date = mdy(Date)) %>%
  filter(Date >= "2021-01-01") %>% 
  select(Date, Variant, value) %>%
  group_by(Variant, Date = floor_date(Date, "month")) %>%
  summarise(value = sum(value))

last_updated <- paste("Source: https://nextstrain.org/sars-cov-2  -  Report Last Updated:",  format(Sys.time(), "%a %b %d %Y %X"))

widget <- hchart(data, "areaspline", hcaes(x = Date, y = value, group = Variant, labels = FALSE)) %>%
  hc_plotOptions(series = list(stacking = 'percent', marker = list(enabled = FALSE))) %>%
  hc_legend(enabled = F) %>%
  hc_add_theme(thm) %>%
  hc_title(text = "India COVID Variants Percentage by Month (PANGO Lineage)", align = "center") %>%
  hc_add_theme(thm) %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE)) %>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
    hc_legend(
    align = "right",
    verticalAlign = "top",
    layout = "vertical"
    ) %>%
  hc_tooltip(shared = TRUE, sort = TRUE, useHTML= TRUE,
              headerFormat = "<b>{point.key}</b><table>",
              pointFormat = "<tr><td style='color: {series.color}'>{series.name}: </td><td style='text-align: right'> {point.percentage:,.2f} %</td></tr>",
             footerFormat = "</table>") %>%
  hc_credits(enabled = TRUE, 
             text = last_updated, 
             position = list(align = "left", x = 10, y = -5))

frameWidget(widget, width = "100%", height = "26rem")

```

`r blogdown::shortcode("alert", "info", "Hover:", "Shows tooltop with percentage of each variant present in data for that month aggregate.")`

This visualization is structured as a percentage chart - where each variant is represented as a percent of total observations for a given month. The tooltip sort order re-shuffles the results to show the most prevalent variants for the month displayed. Even with a relatively short window it's easy to spot two clear trends. B.1.1.7 showed real staying power but has leveled off - and the emergence of B.1.617.1/2 "to the eye" looks like they line up with the spike in cases. __In April those two strains alone make 55.68% of genomically sequenced cases.__ As usual, correlation does not mean causation - and in systems like this even correlation can be elusive, but it's certainly enough to justify spending a bit more time with the data.

### Side Bar: Interest, Concern, Consequence, Variant & Clade

The two references below are worth sharing when looking to grok this domain. The first includes classification of SARS-CoV-2 variants among other information. The second has a concise description of the relationship between variant and clade as well as some notes on the evolution of the SARS-CoV-2 genome.

<div class="row">
  <div class="col-md-6">`r blogdown::shortcode("ext-link-card", "https://www.cdc.gov/coronavirus/2019-ncov/cases-updates/variant-surveillance/variant-info.html")`</div>
  <div class="col-md-6">`r blogdown::shortcode("ext-link-card", "https://www.news-medical.net/health/Viral-Clades-of-SARS-CoV-2.aspx")`</div>
</div>

I've worked at various healthcare-related companies throughout my career, and the one that maps most closely to this domain is `r blogdown::shortcode("url-link", "TherapyEdge", "/vitae/early-years/#TherapyEdge", "_self")`. I gained that formative experience right as AIDS was beginning to be understood as something other than an outright death sentence. So while I don't have the level of knowledge and experience of a clinician or researcher in this field, I know a bit more than the average software engineer. Knowledge is power, and understanding the nuance of how variants can diverge, and how the scientific community responds to and communicates about it will continue to grow in importance as we learn to live with SARS-CoV-2 in the coming years.

## Higher Level View

Looking at clade data from the same observation set shows a more succinct pattern. While it looks like May has some drop off in the partial month data, April shows a clear take-over of 20I/501Y.V1 (primarily due to B.1.1.7) and 21A (due to B.1.617.2) which are the two variants that are currently causing the most worry around the globe.

`r blogdown::shortcode("fancybox-iframe", "/post/COVID-19-India-a-perfect-storm_files/figure-html/widgets/widget_India_COVID_Clades_By_Month.html")`

```{r India_COVID_Clades_By_Month, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

asia <- "/repo/next-strain-ncov/data/asia.csv"

asiadata <- read.csv(asia)

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

data <- select(asiadata, Collection.Data, Country, PANGO.Lineage, Clade) %>%
  filter(Country == "India" 
#         | Country == "Malaysia" 
#         | Country == "Bangladesh" 
#         | Country == "Nepal"
#         | Country == "Sri Lanka"
#         | Country == "Afghanistan"
#         | Country == "Pakistan"
#         | Country == "Maldives"
         ) %>%
  add_column(value = 1) %>%
  rename(Date = Collection.Data) %>%
  mutate(Date = mdy(Date)) %>%
  filter(Date >= "2021-01-01") %>% 
  select(Date, Clade, value) %>%
  group_by(Clade, Date = floor_date(Date, "month")) %>%
  summarise(value = sum(value))

last_updated <- paste("Source: https://nextstrain.org/sars-cov-2  -  Report Last Updated:",  format(Sys.time(), "%a %b %d %Y %X"))

widget <- hchart(data, "areaspline", hcaes(x = Date, y = value, group = Clade, labels = FALSE)) %>%
  hc_plotOptions(series = list(stacking = 'percent', marker = list(enabled = FALSE))) %>%
  hc_legend(enabled = F) %>%
  hc_add_theme(thm) %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE)) %>%
  hc_title(text = "India COVID Clades Percentage by Month (Nextstrain)", align = "center") %>%
  hc_add_theme(thm) %>%
  hc_chart(
    borderColor = 'rgba(160, 160, 160, 0.3)',
    borderRadius = 8,
    borderWidth = 2,
    marginBottom = '80',
    marginTop = '60',
    marginLeft = '60',
    marginRight = '60') %>%
    hc_legend(
    align = "right",
    verticalAlign = "top",
    layout = "vertical"
    ) %>%
  hc_tooltip(shared = TRUE, sort = TRUE, useHTML= TRUE,
              headerFormat = "<b>{point.key}</b><table>",
              pointFormat = "<tr><td style='color: {series.color}'>{series.name}: </td><td style='text-align: right'> {point.percentage:,.2f} %</td></tr>",
             footerFormat = "</table>") %>%
  hc_credits(enabled = TRUE, 
             text = last_updated, 
             position = list(align = "left", x = 10, y = -5))

frameWidget(widget, width = "100%", height = "26rem")

```

`r blogdown::shortcode("alert", "info", "Hover:", "Shows tooltip with percentage of each clade present in data for that month aggregate.")`

Many experts around the globe are watching this very closely, and like variants and clades, so too will the information surrounding them continue to develop.

## News From the World

One of the main concerns surround a potential future "COVID-20" and "COVID-21" is the possibility of those variants reducing the effectiveness of vaccines or available therapies. With that, the contagion would continue to spread, and may eventually overtake once-thought-safe populations. And given the information from those that specialize in this area, certain strains of the virus not only seem more contagious, they also are showing some resistance to certain vaccines and demonstrate potential for accelerated spread.

<div class="row">
  <div class="col-md-6">`r blogdown::shortcode("tweet-single", "1392942636730880003")`</div>
  <div class="col-md-6">`r blogdown::shortcode("tweet-single", "1396332689955950594")`</div>
</div>

During my time at TherapyEdge a significant portion of the company's effort was on determining which strains of the AIDS virus were modulating within a given patient, which would impact the cocktail of ARV and related medications that made an individual's treatment plan. It was like playing whack-a-mole, and if SARS-CoV-2 exhibits similar traits it will be a long time before the healthcare and scientific communities can get in front of it.

## Taking Stock

This isn't the only report I had planned for looking at this data, but given the recent press around `r blogdown::shortcode("url-link", "Modi's COVID response", "https://www.npr.org/2021/05/11/995446333/this-government-has-failed-us-anger-rises-in-india-over-pm-modis-covid-response")`, and the dramatic effect that had on COVID incident rates, I thought some focus time with the data was warranted. Seeing population-leveled incident rates is heart-wrenching enough, but in absolute terms the situation in truly horrific. This isn't *just* a problem for India. It's a problem for *all of us*, and while the scientific community has some clarity on this, I'm hopeful that governments and policy-makers will see beyond their borders and their own personal fortunes to do what's right for the future of our shared world.

<div class="separator">US Health Care Workers Lost to COVID-19</div>
<figure>
<div class="row">
  <div class="col-md-4">`r blogdown::shortcode("tweet-single", COVIDtweets[1])`</div>
  <div class="col-md-4">`r blogdown::shortcode("tweet-single", COVIDtweets[2])`</div>
  <div class="col-md-4">`r blogdown::shortcode("tweet-single", COVIDtweets[3])`</div>
</div>
<figcaption class="caption">The pandemic-related posts on this site are about more than data. <strong>Behind every number is a person, a family and a community.</strong> As reports are refreshed, new selections will also be chosen at random. To see how this is done, <a href="/post/faces-of-covid">see this sidebar</a>.</figcaption>
<hr>
</figure>

`r blogdown::shortcode("alert", "info", "", "If anyone knows a curated Twitter account that lists similar info for Indian healthcare workers I'd be glad to modify my script for this page. I tried a few options using plain text search, but it pulled so many spam posts that were jacking keywords that I couldn't make the switch. This is a global phenomenon and we should honor all of those that have given the last full measure to stem the tide of this horrific pandemic.")`

## The Code Behind the Charts

These charts are similar to many others on this site, with a few wrinkles in display and tooltip styling. I use `r blogdown::shortcode("url-link", "highcharter", "https://jkunst.com/highcharter/")`, the R wrapper for the `r blogdown::shortcode("url-link", "HighCharts", "https://www.highcharts.com")` JavaScript library.

Data for the first visual again uses the Our World In Data record set. Here it's pared down to this year's data and the date field is entitized.

<pre><code class="r">library(highcharter)
library(tidyverse)
library(lubridate)
library(widgetframe)
library(xts)
library(jsonlite)

URL <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/owid-covid-data.csv"

coviddata <- read.csv(URL)

data <- select(coviddata, date, location, new_cases_smoothed_per_million) %>%
  filter(date >= "2021-01-01") %>%
  rename(value = new_cases_smoothed_per_million) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(date = ymd(date)) 

data.wide <-  pivot_wider(data, names_from = location, values_from = value)

data_xts <- xts(data.wide[,-1], order.by = data.wide$date)</code></pre>

Because this visual was about the general shape of the plotted area juxtaposed against the dates I didn't use the "Highstock" visual as I did with Canada and Israel. The largest block of code is the layout of plotted dates, showcasing how the HighCharts wrapper uses a "lists of lists" approach to structuring the hand-off to javascript behind the scenes.

<pre><code class="r">last_updated <- paste("Source: https://ourworldindata.org/  -  Report Last Updated:",  format(Sys.time(), "%a %b %d %Y %X"))

thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

widget <- highchart(type = "chart") %>%
  hc_add_theme(thm) %>%
  hc_title(text = "India's Kumbh Festival Dates Compared To COVID Infection Rates (per Million)", align = "center") %>%
    hc_chart(
      borderColor = 'rgba(160, 160, 160, 0.3)',
      borderRadius = 8,
      borderWidth = 2,
      marginBottom = '80',
      marginTop = '60',
      marginLeft = '60',
      marginRight = '60') %>%
  hc_plotOptions(series = list(marker = list(enabled = FALSE))) %>%
  hc_credits(enabled = TRUE, 
             text = last_updated, 
             position = list(align = "left", x = 10, y = -5)) %>%
  hc_add_series(data_xts$India, 
                type = "area", 
                name = "India", 
                color = "#FF9933", 
                showInLegend = FALSE) %>%
  hc_tooltip(valueDecimals = 2) %>%
  hc_xAxis(type = "datetime",
    plotLines = list(list(
      label = list(text = "First Bath: Maka Sankranti"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-01-14', tz = 'UTC'))
    ),
    list(
      label = list(text = "Second Bath: Mauni Amavasya"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-02-11', tz = 'UTC'))
    ),
    list(
      label = list(text = "Third Bath: Vasant Panchami"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-02-16', tz = 'UTC'))
    ),
    list(
      label = list(text = "Fourth Bath: Ram Navami"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-04-21', tz = 'UTC'))
    ),
    list(
      label = list(text = "First Shahi Bath: Maha Shivratri"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-03-11', tz = 'UTC'))
    ),
    list(
      label = list(text = "Second Shahi Bath: Somvati Amamvasya"),
      color = "#138808",
      width = 3,
      value = datetime_to_timestamp(as.Date('2021-04-12', tz = 'UTC'))
    )
    )) 

frameWidget(widget, width="100%", height="26rem")</code></pre>

This visual that's built from the code below uses a Nextstrain data set, and while that can be generated from their Python/Anaconda project there's a dependency on running the process in a Linux container (as I'm on Windows 10). I've done that kind of work for local dev instances to be eventually be deployed in the Azure stack, but automating a headless Python/Conda process was eating too much time. I'm not sure if I'll go through the process of making it work or if I'll simply re-generate the file a few times per week. Time will tell...

Here you'll see that I looked at countries that border India as well as India itself. I did some data exploration in order to see "fan out" due to other columns showing tracing back to bordering countries. But I'll need to do a more complete job of figuring out how to tie in relevant data, which seemed to be a bridge too far in this exercise. Note that I added a defaulted "1" value column so that roll-ups in the monthly aggregate could then be used by HighCharts to show a late-bound percentage value. (see {point.percentage:,.2f} % in the bottom sample)

<pre><code class="r">library(highcharter)
library(tidyverse)
library(lubridate)
library(widgetframe)
library(jsonlite)

asia <- "/repo/nextstrain-ncov/data/asia.csv"

asiadata <- read.csv(asia)

data <- select(asiadata, Collection.Data, Country, PANGO.Lineage, Clade) %>%
  filter(Country == "India" 
#         | Country == "Malaysia" 
#         | Country == "Bangladesh" 
#         | Country == "Nepal"
#         | Country == "Sri Lanka"
#         | Country == "Afghanistan"
#         | Country == "Pakistan"
#         | Country == "Maldives"
         ) %>%
  add_column(value = 1) %>%
  rename(Date = Collection.Data) %>%
  rename(Variant = PANGO.Lineage) %>%
  mutate(Date = mdy(Date)) %>%
  filter(Date >= "2021-01-01") %>% 
  select(Date, Variant, value) %>%
  group_by(Variant, Date = floor_date(Date, "month")) %>%
  summarise(value = sum(value))</code></pre>

While the area spline and percentage charting makes for a "busy" visual, the key really is the tooltip. Note how HTML for building that mini-table is used to construct a "shared" tool tip, where mousing over any value will show the full range of values for every aggregate for that month. The late-binding of percentage-of-total has a definite "bucket brigade" vibe to it at first glance. But like most of this kind of hybrid/polyglot code exercises it becomes the "new normal" fairly quickly. There's quite a bit of lift that's being carried by the HighCharts "smarts" and that's part of the beauty in the beast.

<pre><code class="r">thm <- hc_theme_merge(
  hc_theme_ffx(),
  hc_theme(title = list(style = list(fontFamily = "Ubuntu")),
           subtitle = list(style = list(fontFamily = "Fira Code")),
           legend = list(itemStyle = list(fontFamily ='Ubuntu'))))

last_updated <- paste("Source: https://nextstrain.org/sars-cov-2  -  Report Last Updated:",  format(Sys.time(), "%a %b %d %Y %X"))

widget <- hchart(data, "areaspline", hcaes(x = Date, y = value, group = Variant, labels = FALSE)) %>%
  hc_plotOptions(series = list(stacking = 'percent', marker = list(enabled = FALSE))) %>%
  hc_legend(enabled = F) %>%
  hc_add_theme(thm) %>%
  hc_title(text = "India COVID Variants Percentage by Month (PANGO Lineage)", align = "center") %>%
  hc_add_theme(thm) %>%
  hc_yAxis(title = list(enabled = FALSE)) %>%
  hc_xAxis(title = list(enabled = FALSE)) %>%
  hc_chart(borderColor = 'rgba(160, 160, 160, 0.3)',
            borderRadius = 8,
            borderWidth = 2,
            marginBottom = '80',
            marginTop = '60',
            marginLeft = '60',
            marginRight = '60') %>%
  hc_legend(align = "right",
            verticalAlign = "top",
            layout = "vertical") %>%
  hc_tooltip(shared = TRUE, sort = TRUE, useHTML= TRUE,
              headerFormat = "&#x3C;b&#x3E;{point.key}&#x3C;/b&#x3E;&#x3C;table&#x3E;",
              pointFormat = "&#x3C;tr&#x3E;&#x3C;td style=&#x27;color: {series.color}&#x27;&#x3E;{series.name}: &#x3C;/td&#x3E;&#x3C;td style=&#x27;text-align: right&#x27;&#x3E; {point.percentage:,.2f} %&#x3C;/td&#x3E;&#x3C;/tr&#x3E;",
              footerFormat = "&#x3C;/table&#x3E;") %>%
  hc_credits(enabled = TRUE, 
              text = last_updated, 
              position = list(align = "left", x = 10, y = -5))

frameWidget(widget, width = "100%", height = "26rem")</code></pre>

I'm seriously considering the lift in "breaking out" the Linux-focused Python tooling for Nextstrain. It's both a way to keep my Python chops relatively sharp as well as lay some groundwork for more exploration of Nextstrain data in F#. Stay tuned!
